Ground state energy of quantum dots using the coupled cluster method

### 2018-06-13 project_2/article/article.tex
Add more discussion in the implementation section

% It's nice to be able to write your own name
\usepackage[T1]{fontenc}
% Automatic clickable links
\usepackage{hyperref}
% SI-units
\usepackage{siunitx}
% Enhanced math formatting
\usepackage{amsmath}
% Extended math symbols
\usepackage{amssymb}
% Include proof environment
\usepackage{amsthm}
\usepackage{enumerate}
% Import the tensor package for tensors
\usepackage{tensor}
% Include dod and dpd fracs
\usepackage{commath}
% Include font for the identity operator
\usepackage{dsfont}
% Include Tikz
\usepackage{tikz}
% Tikz add-on
\usetikzlibrary{shapes,arrows}
\usetikzlibrary{positioning}
% Several figures in the same figure
\usepackage{subfig}
% Appendix
\usepackage[toc, page]{appendix}
\usepackage{simpler-wick}
% Code listings
\usepackage{listings, lstautogobble}
\usepackage{minted}
% Colors
\usepackage{xcolor}
% Directory tree package
\usepackage{dirtree, varwidth}
% Multirow (for table)
\usepackage{multirow}

@@

\section{Theory}
    In this project we will study a system of $N$ interacting electrons. We will
    be looking at a Hamiltonian consisting of a one-body and a two-body part.
    The one-body part is given by
    \begin{align}
        h(\vf{r}_i)
        &= -\half\nabla_i^2 + \half\omega^2 \vf{r}_i^2,
        \label{eq:one_body_hamiltonian}
    \end{align}
    where we use natural units $\hbar = c = e = 1$ and set the mass to unity.
    % Why not write $\hbar = c = e = m = 1$?
    The two-body part is the Coulomb interaction potential,
    \begin{align}
        u(\vf{r}_i, \vf{r}_j)
        &= \frac{1}{\abs{\vf{r}_i - \vf{r}_j}}.
        \label{eq:two_body_operator}
    \end{align}
    We thus get the total Hamiltonian
    \begin{align}
        H &= h + u
        =
        \sum_{i = 1}^N h(\vf{r}_i) + \sum_{i < j}^N u(\vf{r}_i, \vf{r}_j),
    \end{align}
    where $h$ is the full one-body operator and $u$ the full two-body
    operator, i.e., over the entire system.  Working in a basis of $L$ single
    particle functions, $\{\ket{p}\}_{p = 1}^L$. We define the reference Slater
    determinant as
    \begin{align}
        \ket{\Phi_0} &\equiv \ket{1, 2, \dots, N},
    \end{align}
    i.e., a tensorproduct of the $N$ first single particle functions, $\ket{i}$,
    of the system. We call these single particle functions \emph{occupied} as
    they are contained in the Slater determinant.  We will denote the occupied
    indices with $i, j, k, l, \dots \in \{1, \dots, N\}$, the \emph{virtual}
    states with $a, b, c, d, \dots \in \{N + 1, \dots, L\}$ and general indices
    with $p, q, r, s, \dots \in \{1, \dots, L\}$. In terms of sets of basis
    functions we can write this as
    \begin{align}
        \{\ket{p}\}_{p = 1}^L
        = \{\ket{i}\}_{i = 1}^N
        \cup
        \{\ket{a}\}_{a = N + 1}^L,
    \end{align}
    i.e., the general indexed states consists of both occupied and virtual
    states. Note that the single particle functions are orthonormal, i.e.,
    \begin{align}
        \braket{p}{q} = \delta_{pq}.
    \end{align}
    We can construct other Slater determinants in this basis by exciting
    or relaxing the reference determinant. A general excitation is labeled
    $\ket{\Phi_{ij\dots}^{ab\dots}}$ which means that we have removed the single
    particle functions with indices $i, j, \dots$ from the reference and added
    $a, b, \dots$. Note that
    \begin{align}
        \braket{\Phi^{ab\dots}_{ij\dots}}{\Phi_0} = 0,
        \label{eq:excited_overlap}
    \end{align}
    for any excitation.

@@

        \subsubsection{Intermediate computations}
            Looking closely at the tensors in \autoref{eq:CCD_amp_1} one might
            come to realize that many of the terms share a common structure.
            This warrants the search for an algebraic transformation of the
            amplitude equation that has the potential to reduce the amount of
            floating point operations needed to compute it. As it turns out,
            such terms exist and they will decrease the computing time necessary
            by two orders of magnitude for the price of higher storage costs.
            We will define the following "intermediates"\footnote{Note that are
            more than one way of defining the intermediates.},
            \begin{align}
                \label{eq:intermediate1}
                \chi^{ab}_{cd}
                &= \frac{1}{4}t^{ab}_{mn} u^{mn}_{cd}
                + \frac{1}{2}u^{ab}_{cd} \\
                \label{eq:intermediate2}
                \chi^n_j &= \frac{1}{2}t^{cd}_{jm} u^{mn}_{cd} \\
                \label{eq:intermediate3}
                \chi^a_d &= \frac{1}{2} t^{ac}_{nm} u^{nm}_{cd} \\
                \label{eq:intermediate4}
                \chi^{bm}_{jc}
                &= u^{bm}_{jc} + \frac{1}{2}t^{bd}_{jn}u^{mn}_{cd}
            \end{align}
            See Appendix \ref{app:intermediates} for the derivation.  These
            intermediate structures will allow us to rewrite
            \autoref{eq:CCD_amp_1} to,
            \begin{equation}
                \begin{aligned}
                    0 &= u^{ab}_{ij} + f^b_c t^{ac}_{ij}P(ab)
                    - f^k_jt^{ab}_{ik}P(ij) \\
                    &\qquad
                    + t^{cd}_{ij}\chi^{ab}_{cd} + t^{ab}_{in}\chi^n_jP(ij)
                    - t^{bd}_{ij}\chi^a_dP(ab) \\
                    &\qquad
                    + t^{ac}_{im}\chi^{bm}_{jc}P(ab)P(ij)
                    + \frac{1}{2}t^{ab}_{im}u^{mn}_{jn}.
                \end{aligned}
            \end{equation}
            We will discuss the importance of doing this in
            \autoref{subsec:ccd_scaling}.

@@

            The transformations necessary to make all these variations will now
            be contained in $C$. So if we start from a harmonic oscillator (HO)
            basis, $\ket{\alpha}$ we can transform into a better basis, called a
            Hartree-Fock (HF) basis, $\ket{p}$.
            \begin{equation}
                \ket{p}
                = \para{\ket{\alpha}\bra{\alpha}}\ket{p}
                = C^p_\alpha\ket{\alpha}.
            \end{equation}
            We will see how the coupled cluster method with double excitations
            works with both the HO basis and the HF basis, but not equally well.

@@

    \subsection{Program Structure}
        There are three main subsetions within our program structure,
        given in the directory tree below.

@@

        The first subsection, \mintinline{python}{matrix_elemtents},
        contains methods for computing matrix elements in harmonic
        oscillator basis from an analytical
        expression\cite{anisimovas1998energy}.  Computing the matrix
        elements is a very intensive task, and the central functions are
        therefore implemented in C++ with a Cython interface to enable use
        in Python.

        The \mintinline{python}{hartree_fock} subsection contains methods
        for changing from harmonic oscillator basis to Hartree-Fock basis as
        well as an implementation of the self-consistent-field algorithm to
        make this transaction possible.

        The \mintinline{python}{schemes} subsection is arguably the most
        important part of this project. The subsection has three different
        classes that perform the exact same computations, but in different
        and increasingly intelligent ways. First,
        \mintinline{python}{CoupledClusterDoubles} is the most
        straightforward and naïve way to solve the CCD amplitude equations.
        Second, because an overbearing amount of the elements in the
        operator matrices in this problem are zero, we have implemented a
        sparse matrix CDD solver in
        \mintinline{python}{CoupledClusterDoublesSparse}. Third,
        \mintinline{python}{CoupledClusterDoublesOptimized} takes advantage
        of sparse matrices as well, but is also parallellized and optimized
        with memory use and number of floating point operations in mind.

        Both in \mintinline{python}{CoupledClusterDoublesSparse} and
        \mintinline{python}{CoupledClusterDoublesOptimized} we make use
        of the intermediates in equations \ref{eq:intermediate1} to
        \ref{eq:intermediate4}. We have also relied upon implementations
        in C++ for some of the heaviest computations. In the optimized
        class special care was taken to ensure that every contraction
        was computed in an optimal way, with the minimization of memory
        use and CPU time as the overall goal. Moreover, several methods
        in NumPy for doing linear algebra was tested, including but not limited
        to \mintinline{python}{numpy.einsum()}, \mintinline{python}{numpy.tensordot},
        and reshaping datastructures to two dimensions and using
        \mintinline{python}{numpy.dot}. Additionally, we tried to make use of
        the python package \mintinline{python}{dask} which allows for arrays that
        are larger than the computer's memory buffer as well as automated
        concurrency. We found that the dask arrays was a bad fit for our code
        base, but we ended up using \mintinline{python}{numba} for automated
        paralellization of some iterative loops as well as just-in-time
        compialtion of some functions.

    \subsection{Scaling of the coupled cluster doubles method}
        \label{subsec:ccd_scaling}

>>>>

% It's nice to be able to write your own name
\usepackage[T1]{fontenc}
% Automatic clickable links
\usepackage{hyperref}
% SI-units
\usepackage[binary-units]{siunitx}
% Enhanced math formatting
\usepackage{amsmath}
% Extended math symbols
\usepackage{amssymb}
% Include proof environment
\usepackage{amsthm}
\usepackage{enumerate}
% Import the tensor package for tensors
\usepackage{tensor}
% Include dod and dpd fracs
\usepackage{commath}
% Include font for the identity operator
\usepackage{dsfont}
% Include Tikz
\usepackage{tikz}
% Tikz add-on
\usetikzlibrary{shapes,arrows}
\usetikzlibrary{positioning}
% Several figures in the same figure
\usepackage{subfig}
% Appendix
\usepackage[toc, page]{appendix}
\usepackage{simpler-wick}
% Code listings
\usepackage{listings, lstautogobble}
\usepackage{minted}
% Colors
\usepackage{xcolor}
% Directory tree package
\usepackage{dirtree, varwidth}
% Multirow (for table)
\usepackage{multirow}

@@

\section{Theory}
    In this project we will study a system of $N$ interacting electrons. We will
    be looking at a Hamiltonian consisting of a one-body and a two-body part.
    The one-body part is given by
    \begin{align}
        h(\vf{r}_i)
        &= -\half\nabla_i^2 + \half\omega^2 \vf{r}_i^2,
        \label{eq:one_body_hamiltonian}
    \end{align}
    where we use natural units $\hbar = c = e = 1$ and set the mass to unity.
    % Why not write $\hbar = c = e = m = 1$?
    The two-body part is the Coulomb interaction potential,
    \begin{align}
        u(\vf{r}_i, \vf{r}_j)
        &= \frac{1}{\abs{\vf{r}_i - \vf{r}_j}}.
        \label{eq:two_body_operator}
    \end{align}
    We thus get the total Hamiltonian
    \begin{align}
        H &= h + u
        =
        \sum_{i = 1}^N h(\vf{r}_i) + \sum_{i < j}^N u(\vf{r}_i, \vf{r}_j),
    \end{align}
    where $h$ is the full one-body operator and $u$ the full two-body
    operator, i.e., over the entire system.  Working in a basis of $L$ single
    particle functions, $\{\ket{p}\}_{p = 1}^L$. We define the reference Slater
    determinant as
    \begin{align}
        \ket{\Phi_0} &\equiv \ket{1, 2, \dots, N},
    \end{align}
    i.e., a tensorproduct of the $N$ first single particle functions, $\ket{i}$,
    of the system. We call these single particle functions \emph{occupied} as
    they are contained in the Slater determinant.  We will denote the occupied
    indices with $i, j, k, l, \dots \in \{1, \dots, N\}$, the \emph{virtual}
    states with $a, b, c, d, \dots \in \{N + 1, \dots, L\}$ and general indices
    with $p, q, r, s, \dots \in \{1, \dots, L\}$. We define $M = L - N$ as the
    number of virtual states. In terms of sets of basis functions we can write
    this as
    \begin{align}
        \{\ket{p}\}_{p = 1}^L
        = \{\ket{i}\}_{i = 1}^N
        \cup
        \{\ket{a}\}_{a = N + 1}^L,
    \end{align}
    i.e., the general indexed states consists of both occupied and virtual
    states. Note that the single particle functions are orthonormal, i.e.,
    \begin{align}
        \braket{p}{q} = \delta_{pq}.
    \end{align}
    We can construct other Slater determinants in this basis by exciting
    or relaxing the reference determinant. A general excitation is labeled
    $\ket{\Phi_{ij\dots}^{ab\dots}}$ which means that we have removed the single
    particle functions with indices $i, j, \dots$ from the reference and added
    $a, b, \dots$. Note that
    \begin{align}
        \braket{\Phi^{ab\dots}_{ij\dots}}{\Phi_0} = 0,
        \label{eq:excited_overlap}
    \end{align}
    for any excitation.

@@

        \subsubsection{Intermediate computations}
            Looking closely at the tensors in \autoref{eq:CCD_amp_1} one might
            come to realize that many of the terms share a common structure.
            This warrants the search for an algebraic transformation of the
            amplitude equation that has the potential to reduce the amount of
            floating point operations needed to compute it. As it turns out,
            such terms exist and they will decrease the computing time necessary
            by two orders of magnitude for the price of higher storage costs.
            We will define the following "intermediates"\footnote{Note that are
            more than one way of defining the intermediates.},
            \begin{align}
                \label{eq:intermediate1}
                \chi^{ab}_{cd}
                &= \frac{1}{4}t^{ab}_{mn} u^{mn}_{cd}
                + \frac{1}{2}u^{ab}_{cd} \\
                \label{eq:intermediate2}
                \chi^n_j &= \frac{1}{2}t^{cd}_{jm} u^{mn}_{cd} \\
                \label{eq:intermediate3}
                \chi^a_d &= \frac{1}{2} t^{ac}_{nm} u^{nm}_{cd} \\
                \label{eq:intermediate4}
                \chi^{bm}_{jc}
                &= u^{bm}_{jc} + \frac{1}{2}t^{bd}_{jn}u^{mn}_{cd}
            \end{align}
            See Appendix \ref{app:intermediates} for the derivation.  These
            intermediate structures will allow us to rewrite
            \autoref{eq:CCD_amp_1} to,
            \begin{equation}
                \begin{aligned}
                    0 &= u^{ab}_{ij} + f^b_c t^{ac}_{ij}P(ab)
                    - f^k_jt^{ab}_{ik}P(ij) \\
                    &\qquad
                    + t^{cd}_{ij}\chi^{ab}_{cd} + t^{ab}_{in}\chi^n_jP(ij)
                    - t^{bd}_{ij}\chi^a_dP(ab) \\
                    &\qquad
                    + t^{ac}_{im}\chi^{bm}_{jc}P(ab)P(ij)
                    + \frac{1}{2}t^{ab}_{im}u^{mn}_{jn}.
                \end{aligned}
            \end{equation}
            Doing this reduces the number of FLOPS from $\mathcal{O}(M^4N^4)$
            for the heaviest contraction, i.e., the contraction
            \begin{align}
                g(f, u, t) \gets
                \frac{1}{4}t^{cd}_{ij}t^{ab}_{nm}u^{mn}_{cd},
            \end{align}
            to $\mathcal{O}(M^4N^2)$ when using the largest intermediate to do
            the contraction
            \begin{align}
                g(f, u, t) \gets t^{cd}_{ij}\chi^{ab}_{cd},
            \end{align}
            for the price of storing $M^4$ more matrix elements and
            pre-computing the intermediate.

@@

            The transformations necessary to make all these variations will now
            be contained in $C$. So if we start from a harmonic oscillator (HO)
            basis, $\ket{\alpha}$ we can transform into a better basis, called a
            Hartree-Fock (HF) basis, $\ket{p}$.
            \begin{equation}
                \ket{p}
                = \para{\ket{\alpha}\bra{\alpha}}\ket{p}
                = C^{\alpha}_p\ket{\alpha}.
            \end{equation}
            We will see how the coupled cluster method with double excitations
            works with both the HO basis and the HF basis, but not equally well.

@@

    \subsection{Program Structure}
        There are three main subsections within our program structure, given in
        the directory tree below.

@@

        The first subsection, \mintinline{bash}{matrix_elemtents},
        contains methods for computing matrix elements in harmonic
        oscillator basis from an analytical
        expression\cite{anisimovas1998energy}.  Computing the matrix
        elements is a very intensive task, and the central functions are
        therefore implemented in C++ with a Cython interface to enable use in
        Python\footnote{The C++ code is inherited from Alocias Mariadason from
        his project in FYS4411 last year with slight modifications. The original
        source code can be found here
        \url{https://github.com/Oo1Insane1oO/FYS4411/blob/master/project1/src/Coulomb_Functions.cpp}
        and
        \url{https://github.com/Oo1Insane1oO/FYS4411/blob/master/project1/src/Coulomb_Functions.hpp}.
        }.

        The \mintinline{bash}{hartree_fock} subsection contains methods
        for changing from harmonic oscillator basis to Hartree-Fock basis as
        well as an implementation of the self-consistent-field algorithm to
        make this transaction possible.

        The \mintinline{bash}{schemes} subsection is arguably the most important
        part of this project. The subsection has three different classes that
        perform the exact same computations, but in different and increasingly
        intelligent ways. First, \mintinline{python}{CoupledClusterDoubles} is
        the most straightforward and naïve way to solve the CCD amplitude
        equations.  Second, because an overbearing amount of the elements in the
        operator matrices in this problem are zero, we have implemented a sparse
        matrix CCD solver in \mintinline{python}{CoupledClusterDoublesSparse}.
        Third, \mintinline{python}{CoupledClusterDoublesOptimized} is
        parallellized and optimized with memory use and number of floating point
        operations in mind.

    \subsection{Index mapping}
        The index mapping generating a $(n, m) \mapsto p$ and $p \mapsto (n, m)$
        is a Python adaption of the code found here:
        \url{https://github.com/ManyBodyPhysics/LectureNotesPhysics/blob/master/Programs/Chapter8-programs/cpp/CCD/src/qDots2DSPBasis.cpp#L18-L68}.

    \subsection{Generating the antisymmetric two-body elements}
        Constructing the Coulomb elements is by far the most intensive task
        that is done. Luckily, we only need to do this once for each basis size
        as the elements do not change and can be reused. We therefore compute
        the elements once and store all the non-zero elements as Python
        pickle-objects. If we wish to do a new run we first look for the
        relevant Coulomb elements in the \texttt{dat}-directory before
        attempting to create them anew.

        After the elements are created/read we have to scale them with the
        current frequency, i.e.,
        \begin{align}
            \bra{\alpha\beta}u\ket{\gamma\delta}
            = \sqrt{\omega}\bra{12}u\ket{34},
        \end{align}
        where $\bra{12}u\ket{34}$ is found from E. Anisimovas and A.  Matulis
        (Equation A2)\cite{anisimovas1998energy}. Next we antisymmetrize the
        two-body elements and include spin which doubles each dimension of the
        orbital integrals.
        \begin{align}
            u^{\alpha\beta}_{\gamma\delta}
            &=
            \delta_{\sigma_{\alpha}\sigma_{\gamma}}
            \delta_{\sigma_{\beta}\sigma_{\delta}}
            \bra{\alpha\beta}u\ket{\gamma\delta}
            \nonumber \\
            &\qquad
            -
            \delta_{\sigma_{\alpha}\sigma_{\delta}}
            \delta_{\sigma_{\beta}\sigma_{\gamma}}
            \bra{\alpha\beta}u\ket{\delta\gamma}.
        \end{align}
        The antisymmetrization process uses four nested loops and is therefore
        implemented in Cython to reduce the computation time.

    \subsection{Reducing the complexity of the basis transformation}
        Having computed the RHF energy and the basis transformation matrix $C$
        we can transform from the HO basis to the HF basis. For the one-body
        matrix elements this becomes
        \begin{align}
            \bra{p}h\ket{q}
            &= {C^p_{\alpha}}^{\dagger}C^{\beta}_q\bra{\alpha}h\ket{\beta}.
        \end{align}
        Note the usage of the Hermitian conjugate and complex conjugate. We
        define them in such a manner
        \begin{align}
            {C^p_{\alpha}}^{\dagger}
            = {C^{\alpha}_p}^{*}.
        \end{align}
        For the two-body elements this is a more complicated
        task\footnote{Imagine the poor fellows that needs to deal with
        three-body elements...}.
        \begin{align}
            \bra{pq}u\ket{rs}
            &=
            {C^p_{\alpha}}^{\dagger}{C^q_{\beta}}^{\dagger}
            C^{\gamma}_r C^{\delta}_s
            \bra{\alpha\beta}u\ket{\gamma\delta},
        \end{align}
        which translates into eight nested loops when programming the
        expression as it stands. Luckily, by noticing that no indices are
        repeated more than twice we can do each loop separately viz.
        \begin{align}
            \bra{pq}u\ket{rs}
            &=
            {C^p_{\alpha}}^{\dagger}
            \brak{
                {C^q_{\beta}}^{\dagger}
                \brak{
                    C^{\gamma}_r
                    \brak{
                        C^{\delta}_s
                        \bra{\alpha\beta}u\ket{\gamma\delta}
                    }
                }
            }.
        \end{align}
        This reduces the computations to four quintuply nested
        loops\footnote{This trick was made aware to us through
        \url{http://sirius.chem.vt.edu/wiki/doku.php?id=crawdad:programming:project4}.
        }.

    \subsection{Implementation details}
        Both in \mintinline{python}{CoupledClusterDoublesSparse} and
        \mintinline{python}{CoupledClusterDoublesOptimized} we make use of the
        intermediates in equations \ref{eq:intermediate1} to
        \ref{eq:intermediate4}.  In the optimized class special care was taken
        to ensure that every contraction was computed in an optimal way, with
        the minimization of memory use and CPU time as the overall goal.
        Moreover, several methods in NumPy for doing linear algebra was tested,
        including but not limited to \mintinline{python}{numpy.einsum},
        \mintinline{python}{numpy.tensordot}, and reshaping datastructures to
        two dimensions and using \mintinline{python}{numpy.matmul}.
        Additionally, we tried to make use of the python package
        \mintinline{python}{dask} which allows for arrays that are larger than
        the computer's memory buffer as well as automated concurrency. We found
        that the dask arrays was a bad fit for our code base, but we ended up
        using \mintinline{python}{numba} for paralellization of some
        iterative loops as well as just-in-time compilation of some functions.

    \subsection{Scaling of the coupled cluster doubles method}
        When talking about the scaling of the coupled cluster method there are
        two aspects to consider.
        \begin{enumerate}
            \item The number of FLOPS to compute $g(f, u, t)$.
            \item The number of bytes needed to store the arrays of matrix
                elements.
        \end{enumerate}
        For the CCD method the naïve solution using the form of $g(f, u, t)$
        without the intermediates results in $\mathcal{O}(M^4N^4)$ FLOPS each
        iteration. Using the form of the intermediates we have set up this
        reduces to $\mathcal{O}(M^4N^2)$, i.e., two order of magnitude less. The
        effect of this is \emph{dramatic} and will be shown in the results
        section in due time.

        A downside of using the intermediates\footnote{Particularly our form for
        the intermediates as we store $\chi^{ab}_{cd}$ instead of a rewriting
        this intermediate as $\chi^{ij}_{kl}$ which is also possible.} is the
        higher storage costs. Lucky enough for us, the largest storage cost
        comes from storing $u^{pq}_{rs}$ and overshadows all other terms as it
        scales as $L^4$. See \autoref{tab:memory_u} for some example storage
        costs. For a normal desktop computer we quickly reach a threshold where
        $u^{pq}_{rs}$ will eat up all memory and activate disk storage thus
        triggering a massive slowdown in computational speed.

        \begin{table}
            \centering
            \caption{Storage costs of $u^{pq}_{rs}$ in gigabytes for a given
            shell $R$ and the corresponding number of basis functions $L$
            assuming $\SI{8}{\byte}$ (\texttt{double}) precision.}
            \begin{ruledtabular}
                \begin{tabular}{ccc}
                    $R$ & $L$ & $u^{pq}_{rs}$ $[\si{\giga\byte}]$ \\
                    \hline
                    $8$ & $72$ & $0.2$ \\
                    $9$ & $90$ & $0.49$ \\
                    $10$ & $110$ & $1.1$ \\
                    $11$ & $132$ & $2.3$ \\
                    $12$ & $156$ & $4.4$ \\
                    $13$ & $182$ & $8.2$ \\
                    $14$ & $210$ & $14.5$
                \end{tabular}
            \end{ruledtabular}
            \label{tab:memory_u}
        \end{table}

        For larger shell values we have to be particularly careful in terms of
        the memory usage. This can be semi-problematic when using a programming
        language such as Python as memory usage is camouflaged. Using NumPy
        allows us to reuse memory buffers quite efficiently, but we have to be
        careful what type of operations we do and if they incur copying.

        \subsubsection{Sparse implementation}
            It turns out that most of the elements in both
            $u^{\alpha\beta}_{\gamma\delta}$ (the HO basis) and $u^{pq}_{rs}$
            (the HF basis) are zero. Roughly $0.014\%$ and $0.082\%$ of the
            elements are non-zero respectively (for $R = 12$). This inspired us
            to look for a way to do sparse CCD. Using the
            \texttt{sparse}-library in Python\footnote{Found here:
            \url{https://github.com/pydata/sparse}.} we were able to set up a
            sparse implementation. Unfortunately being efficient in terms of
            re-using memory is quite complicated and there are definitely
            improvements to be made. Nonetheless the performance of the method
            is comparable to the optimized code.

        \subsubsection{Parallelizing the contractions}
            The easiest way to parallelize the coupled cluster code is to use
            OpenMP on the contractions. In particular the heavy contractions. We
            solved this using Numba with the \texttt{prange} function which
            essentially translates into an OpenMP directive. This turned out to
            be slower than the built-in NumPy calls, especially since NumPy
            comes with the option of running in parallel if the code is built
            with BLAS by setting the environment variable
            \texttt{OMP\_NUM\_THREADS}. For example:
            \begin{bash}
                export OMP_NUM_THREADS=4
            \end{bash}

            Therefore, our most optimized code uses the matrix multiplication
            code (\texttt{np.matmul}) from NumPy which translates into the BLAS
            call \texttt{DGEMM} for matrix-matrix multiplication. This requires
            us to reshape the four-dimensional tensors as matrices and write
            them in such a way that we can do the contraction using matrix
            multiplication. Here NumPy comes to our rescue as a reshape is done
            inplace and all memory is stored as a contiguous array.
